---
title: "Data Science Summer School - MSR NYC"
author: "Nikola Baci and Yasiris Ortiz"
date: "6/21/2021"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup}
suppressMessages(library(tidyverse))
suppressMessages(library(lubridate))
suppressMessages(library(gganimate))
```

## Small share of US police draw third of complaints in big cities
#### Article by [FINANCIAL TIMES](https://www.ft.com/content/141182fc-7727-4af8-a555-5418fa46d09e)

## Overview of the article
The [article](https://www.ft.com/content/141182fc-7727-4af8-a555-5418fa46d09e) published
by Financial Times on May 28, 2021, one year after the death of George Floyd, 
tries to analyse the complaints that are filed against police officers in three major cities:
NYC, Philadelphia, Chicago. The article points out that the ten percent of the officer with
the most complaints received a third of all complaints during the period 2007-2017 for NYC and 
Chicago and 2015-2021 for Philly. 

In this notebook, we are trying to replicate these finding and achieve the same bar graphs as
article. In addition, we expand the notebook with further analysis such as the gender and the
race of the officers that received the complaints.

## Data

Data for NYC can be found [here](https://github.com/new-york-civil-liberties-union/NYPD-Misconduct-Complaint-Database-Updated).

Data for Chicago can be found [1](https://github.com/invinst/chicago-police-data/blob/master/data/unified_data/complaints/complaints-accused.csv.gz), [2](https://github.com/invinst/chicago-police-data/blob/master/data/unified_data/complaints/complaints-complaints.csv.gz), [3](https://github.com/invinst/chicago-police-data/blob/master/data/unified_data/complaints/officer-filed-complaints__2017-09.csv.gz) and
[4](https://github.com/invinst/chicago-police-data/blob/master/data/unified_data/complaints/complaints-complainants.csv.gz).

Data for Philadelphia can be found [1](https://github.com/sdl60660/philly_police_complaints/blob/master/raw_data/ppd_complaints.csv), [2](https://github.com/sdl60660/philly_police_complaints/blob/master/raw_data/ppd_complaints_7-18.csv), [3](https://github.com/sdl60660/philly_police_complaints/blob/master/raw_data/ppd_complaint_disciplines.csv), [4](https://www.opendataphilly.org/dataset/police-complaints/resource/e7477284-0045-4f37-8aeb-182616f736e8) and [5](https://www.opendataphilly.org/dataset/police-complaints/resource/7f7d472f-c49c-4364-b6e0-3a079e6b7d7f).

```{r Data}
nyc_data <- read.csv("data/nyc_misconduct_complaint_dataset.csv")

chicago_complaints_accused <- read.csv("data/chicago_complaints_accused.csv")

chicago_complaints_complaints <- read.csv("data/chicago_complaints_complaints.csv")

chicago_complaints_by_officers <- read.csv("data/chicago_officer_filed_complaints.csv")

philly_open_data_1 <- read.csv('data/ppd_complaints_open_philly.csv',header=T) 

philly_open_data_2 <- read.csv('data/ppd_complaint_disciplines_open_philly.csv',header=T)

philly_sam_1 <- read.csv('data/philly_complaints_dataset_sam_cap_number.csv', header = T) 

philly_sam_2 <- read.csv('data/philly_complaints_dataset_sam.csv',header=T) 

philly_sam_3 <- read.csv('data/philly_complaints_disciplines_sam.csv') 

```

## New York City

Replication of NYC bar graph from the article. We first create the necessary
features and tables, then we proceed to graphing.

The resulting graph looks identical to the article. 

### NYC Data Wrangling


```{r NYC-Data-Wrangling}
nyc_data_original <- nyc_data

nyc_data <- nyc_data %>%
  mutate(ReceivedDate = mdy(ReceivedDate)) %>% #convert from char to Date type
  filter(ReceivedDate >= "2007-01-01", ReceivedDate <= "2017-12-31") #get only the period between 2007-2017
```

### NYC Feature Engineering

```{r NYC-Featuer-Engineering}

# Goal: Bucket each officer in one of the 10 buckets by the number of complaints.
# Note: Some complaints include several allegations for the same officer
# or for a number of officers
# Note: The names are not unique or sufficient to be used to identify each officer.

nyc_by_num_complaints <- nyc_data %>%
  group_by(OfficerID, ComplaintID) %>%
  summarize() %>%                                       # count one officer per complaint 
  summarize(num_of_complaints = n()) %>%                # count the total complaints per officer
  arrange(num_of_complaints)                            # order in ascending order

total_complaints <- sum(nyc_by_num_complaints$num_of_complaints)  # sum of all complaints, used to obtain the percents for each bin

nyc_by_num_complaints <- nyc_by_num_complaints %>%
  mutate(bin_number = 1:nrow(nyc_by_num_complaints)) %>%  # create a new column to place the bin number 1 to 10
  mutate(bin_number = ntile(bin_number, 10)) # use the ntile function to put rows in the right bins

# collapse the table to only 10 rows that represent the bins
# for each bin find the percentage of the complaints that the bin has
# compared to the total
nyc_by_num_complaints <- nyc_by_num_complaints %>%
  group_by(bin_number) %>%
  summarize(total_complaints = (sum(num_of_complaints) / total_complaints) * 100)
```
### NYC Graphing

```{r NYC-Graph, fig.align = 'center'}
nyc_by_num_complaints

nyc_by_num_complaints <-  nyc_by_num_complaints %>%
  mutate(color = c(rep(0, 9), 1)) # to help with the bar colors

# graph the deciles (bins) and their percentages
nyc_by_num_complaints %>%
  ggplot()+
  geom_histogram(aes(x = bin_number, y = total_complaints, fill = color), stat="identity", show.legend = F)+
  scale_x_continuous(breaks = seq(1, 10, 1)) + #set x-axis to go up by 1
  xlab("Decile")+
  ylab("Percent of total complaints") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"), # set background to white
          panel.grid.major.y = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray")) # only horizontal marks
  
```

## Chicago 

### Chicago Data Wrangling
Now we proceed to working with the Chicago data and obtaining the tables to
draw the bars.

```{r Chicago-Data-Wrangling}

chicago_accused_selected <-  chicago_complaints_accused %>%
  select(UID, cr_id, link_UID) #select only 3 out of the 10 columns

chicago_complaints_selected <- chicago_complaints_complaints %>%
  select(cr_id, complaint_date, closed_date) %>% #select only 3 out of the 11 columns
  mutate(complaint_date = ymd(complaint_date), closed_date = ymd(closed_date)) #turn char dates to Date-types

chicago_officers_selected <- chicago_complaints_by_officers %>%
  mutate(cr_id = as.character(cr_id)) #add a flag column, will help in keeping only the citizens complaints

#create a table that has three essential columns: officer id, complaint id, and data of complaint
chicago_complaints <- inner_join(chicago_accused_selected, chicago_complaints_selected, by="cr_id") %>%
  filter(complaint_date >= "2007-01-01",  closed_date <= "2017-12-31") #choose only period 2007-2017

#remove complaints that are done by officers or that are not closed yet
chicago_complaints_citizens <- anti_join(chicago_complaints, chicago_officers_selected, by="cr_id") %>%
   filter(!is.na(closed_date))

```
### Chicago Feature Engineering

```{r Chicago-Feature-Engineering}
#Note: the code is the same as in NYC-Graph above, the only thing that changes is the
#data table in the next line and the variables to indicate that this data is for Chicago.
chicago_by_num_complaints <- chicago_complaints_citizens %>%
  group_by(UID, cr_id) %>%
  summarize() %>%
  summarize(num_of_complaints = n()) %>% 
  arrange(num_of_complaints)

total_complaints <- sum(chicago_by_num_complaints$num_of_complaints)  

chicago_by_num_complaints <- chicago_by_num_complaints %>%
  mutate(bin_number = 1:nrow(chicago_by_num_complaints)) %>%  
  mutate(bin_number = ntile(bin_number, 10))

chicago_by_num_complaints <- chicago_by_num_complaints %>%
  group_by(bin_number) %>%
  summarize(total_complaints = (sum(num_of_complaints) / total_complaints) * 100)

```

### Chicago Graphing

Now we can graph the decile bars. 
Note: the data includes the only the citizen complaints that were registered on or after
01/01/2007 and were closed before 01/01/2018.

```{r Explore Chicago-Graph, fig.align = 'center'}

chicago_by_num_complaints

chicago_by_num_complaints <-  chicago_by_num_complaints %>%
  mutate(color = c(rep(0, 9), 1))

chicago_by_num_complaints %>%
  ggplot()+
  geom_histogram(aes(x = bin_number, y = total_complaints, fill = color), stat="identity", show.legend = F)+
  scale_x_continuous(breaks = seq(1, 10, 1))+
  xlab("Decile")+
  ylab("Percent of total complaints") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())
```

## Philadelphia

### Philly Data Wrangling

```{r Philly-Data-Wrangling}

philly_open_data_1 <- philly_open_data_1 %>%
  subset(select=-summary) %>%
  mutate(date_received = as.Date(date_received))

philly_open_data_join <- left_join(philly_open_data_2, philly_open_data_1, by='complaint_id') %>%
  filter(date_received >= '2015-04-01')

philly_sam_1 <- philly_sam_1 %>%
  mutate(date_received = as.Date(date_received))
philly_sam_1 <-  rename(philly_sam_1, complaint_id = cap_number, district_occurrence = dist_occurrence)


philly_sam_2 <- philly_sam_2 %>% 
  mutate(date_received = mdy(date_received)) %>% # change the date format here 
  select(-shortened_summary)


philly_sam_3 <-  philly_sam_3%>%
  select(complaint_id, officer_id)

philly_data <- bind_rows(philly_sam_1, philly_sam_2) %>%
  left_join(philly_sam_3, by="complaint_id") %>%
  filter(date_received >= "2015-01-01", ! is.na(officer_id))

```

### Philly Feature Engineering

```{r Philly-Feature-Engineering}

philly_by_num_complaints <- philly_data %>%
  group_by(officer_id, complaint_id) %>%
  summarize() %>%                                       
  summarize(num_of_complaints = n()) %>%                
  arrange(num_of_complaints)                            

total_complaints <- sum(philly_by_num_complaints$num_of_complaints)  
philly_by_num_complaints <- philly_by_num_complaints %>%
  mutate(bin_number = 1:nrow(philly_by_num_complaints)) %>%  
  mutate(bin_number = ntile(bin_number, 10)) 

philly_by_num_complaints <- philly_by_num_complaints %>%
  group_by(bin_number) %>%
  summarize(total_complaints = (sum(num_of_complaints) / total_complaints) * 100)
```

### Philly Graphing

```{r Philly-Graph, fig.align = 'center'}

philly_by_num_complaints

philly_by_num_complaints <-  philly_by_num_complaints %>%
  mutate(color = c(rep(0, 9), 1))

philly_by_num_complaints %>%
  ggplot()+
  geom_histogram(aes(x = bin_number, y = total_complaints, fill = color), stat="identity", show.legend = F)+
  scale_x_continuous(breaks = seq(1, 10, 1))+
  scale_y_continuous(breaks = seq(0, 30, 5))+
  xlab("Decile")+
  ylab("Percent of total complaints") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())
```

## Extended Analysis

### NYC Complaints by Officer Gender

In this section we analyze the complaints by gender. The first graph shows
the distribution of all complaints and the second graph shows the distribution
of the 10th decile only (the officers with top 10% of complaints).
```{r NYC-Complaints-Gender, fig.align = 'center'}

nyc_complaint_by_gender <- nyc_data %>%
  group_by(OfficerID, ComplaintID, OfficerGender) %>%
  suppressMessages(summarize()) %>% 
  ungroup() %>%
  group_by(OfficerID) %>%
  summarize(num_of_complaints = n(), OfficerGender) %>%
  ungroup()
  
#complaints by gender among all complaints
total_complaints <- sum(nyc_complaint_by_gender$num_of_complaints)
nyc_complaint_by_gender %>%
  group_by(OfficerGender) %>%
  summarize(percent = sum(num_of_complaints) / total_complaints) %>% 
  ggplot() +
  geom_bar(aes(x = OfficerGender, y = percent), stat = "identity", show.legend = F, fill = "brown") +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  xlab("Officer Gender") +
  ylab("Propotion of Complaints") +
  labs(title = "Total Complaints by Officer Gender") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())

#complaints by gender only bin 10
nyc_complaint_by_gender <- nyc_complaint_by_gender %>%
  mutate(bin_number = 1:nrow(nyc_complaint_by_gender)) %>% # create a new column to place the bin number 1 to 10
  mutate(bin_number = ntile(bin_number, 10)) %>%
  filter(bin_number == 10)

bin_10_total_complaints <- sum(nyc_complaint_by_gender$num_of_complaints)

nyc_complaint_by_gender %>%
  group_by(OfficerGender) %>%
  summarize(percent = sum(num_of_complaints) / bin_10_total_complaints) %>% 
  ggplot() +
  geom_bar(aes(x = OfficerGender, y = percent), stat = "identity", show.legend = F,  fill = "brown") +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  xlab("Officer Gender") +
  ylab("Propotion of Complaints") +
  labs(title = "Top 10% of Complaints by Officer Gender") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())
```

### Complaints by Officer Race

In this section we will analyze the race of the officers and what proportion of
the complaints each race receives. 
Note: Please keep in mind that the we need more data to reach a conclusion. Things 
to keep in mind is the distribution of race in the police force as well as population
of these cities. 
Hypothetical example, if NYC population is composed of 80% white and 
the police force has the same proportion of white officers, then it is more likely that
the below bars will show an overwhelming amount of complaints towards white officers.

The results show that more than 50% of all complaints in NYC and Chicago go towards white officers
#### NYC 

```{r NYC-Complaints-By-Officer-Race, fig.align = 'center'}
nyc_complaints_by_officer_race <- nyc_data %>% 
  group_by(OfficerID, ComplaintID, OfficerRace) %>%
  suppressMessages(summarize()) %>%
  ungroup()

nyc_complaints_by_officer_race <- nyc_complaints_by_officer_race %>%
  group_by(OfficerRace) %>%
  summarize(count = n()) %>%
  mutate(percent = count / nrow(nyc_complaints_by_officer_race) * 100)

nyc_complaints_by_officer_race

nyc_complaints_by_officer_race %>%
  ggplot() +
  geom_bar(aes(x= OfficerRace, y = percent), stat = "identity", fill = "brown") +
  labs(title = "New York 2007-2017 complaints by officer race") +
  xlab("Officer Race") +
  ylab("Percent") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())
```

The animation below shows the changes of the complaints that officers
received, separated by their race, over the period 2007-2017.

```{r NYC-Complaint-Animation-Over-Years, fig.align = 'center'}

nyc_complaints_timeline <-  nyc_data %>%
  mutate(year = year(ReceivedDate)) %>%
  group_by(year, OfficerRace) %>%
  summarize(count = n())

plot_transition <- ggplot(data = nyc_complaints_timeline) +
  geom_bar(aes(x = OfficerRace, y = count), stat = "identity", fill = "brown") +
  xlab("Officer Race") +
  ylab("Number of Complaints") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank(),
        plot.title=element_text(size=25, hjust=0.5, face="bold", colour="grey", vjust=-1)) +
  transition_states(year, transition_length = 10, state_length = 1)+
  labs(title = 'Year : {closest_state}')
  
animate(plot_transition, fps=10)
```

#### Chicago

```{r Chicago-Complaints-By-Officer-Race, fig.align = 'center'}
chicago_officer_profiles <- read.csv("data/chicago_officer_profile.csv")
chicago_officer_profiles <- chicago_officer_profiles %>%
  select(UID, race)

chicago_complaints_and_officer <- left_join(chicago_complaints, chicago_officer_profiles, by="UID") %>%
  filter(race != "") #there is one officer whose race is unknown

chicago_complaints_by_offcier_race <- chicago_complaints_and_officer %>%
  group_by(UID, cr_id, race) %>%
  suppressMessages(summarize()) %>%
  ungroup()

chicago_complaints_by_offcier_race <- chicago_complaints_by_offcier_race %>%
  group_by(race) %>%
  summarize(count = n()) %>%
  mutate(percent = count / nrow(chicago_complaints_by_offcier_race) * 100) %>%
  mutate(race = factor(race)) %>%
  arrange(percent)

chicago_complaints_by_offcier_race %>%
  ggplot() +
  geom_bar(aes(x = fct_reorder(race, percent), y = percent), stat = "identity", fill = "brown") +
  labs(title = "Chicago 2007-2017 complaints by officer race") +
  xlab("Officer Race") +
  ylab("Percent") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())
```

### Distribution of Complaints by Officer Race

In this section we separate officers by their race and observer the count of
officers by the number of complaints they have.

#### NYC
```{r NYC-Complaint-Distribution-By-Race, fig.align = 'center'}
nyc_complaints_distribution_by_officer_race <- nyc_data %>% 
  group_by(OfficerID, ComplaintID, OfficerRace) %>%
  suppressMessages(summarize()) %>%
  ungroup

nyc_complaints_distribution_by_officer_race <-  nyc_complaints_distribution_by_officer_race %>%
  group_by(OfficerID, OfficerRace) %>%
  summarize(num_of_complaints = n())

nyc_complaints_distribution_by_officer_race %>%
  ggplot() +
  geom_histogram(aes(x = num_of_complaints), fill = "brown") +
  labs(title = "NYC Complaint Distribution by Race") +
  xlab("Number of Complaints") +
  ylab("Number of Officers with X Complaints") +
  facet_wrap(~ OfficerRace, scale = "free_y") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())

nyc_complaints_distribution_by_officer_race %>%
  group_by(OfficerRace, num_of_complaints) %>%
  summarize(count = n()) %>%
  ggplot() +
  geom_jitter(aes(x = num_of_complaints, y = count), color = "brown") +
  labs(title = "NYC Complaint Distribution by Officer's Race") +
  xlab("Number of Complaints") +
  ylab("Number of Officers with X Complaints") +
  facet_wrap(~ OfficerRace, scales = "free_y") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())



```

#### Chicago

```{r Chicago-Complaint-Distribution-By-Race, fig.align = 'center'}

chicago_complaints_distribution_by_officer_race <- chicago_complaints_and_officer %>% 
  group_by(UID, cr_id, race) %>%
  summarize() %>%
  ungroup

chicago_complaints_distribution_by_officer_race <-  chicago_complaints_distribution_by_officer_race %>%
  group_by(UID, race) %>%
  summarize(num_of_complaints = n())

chicago_complaints_distribution_by_officer_race %>%
  ggplot() +
  geom_histogram(aes(x = num_of_complaints), fill = "brown") +
  labs(title = "Chicago Complaint Distribution by Officer's Race") +
  xlab("Number of Complaints") +
  ylab("Number of Officers with X Complaints") +
  facet_wrap(~ race, scales = "free_y") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())

chicago_complaints_distribution_by_officer_race %>%
  group_by(race, num_of_complaints) %>%
  summarize(count = n()) %>%
  ggplot() +
  geom_jitter(aes(x = num_of_complaints, y = count), color = "brown") +
  labs(title = "Chicago Complaint Distribution by Race") +
  xlab("Number of Complaints") +
  ylab("Number of Officers with X Complaints") +
  facet_wrap(~ race, scales = "free_y") +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())



```

### Distribution of Impacted Race by Officer Race
Lastly, we try to understand the distribution of the citizens that filed a complaint in 
NYC and Chicago during the 2007-2017 period.

Results show that the black community is filing more complaints.

#### NYC
```{r NYC-Distribution-of-Complainants-by-Officer-Race, fig.align = 'center'}
nyc_complaniant <- nyc_data %>%
  filter(ImpactedRace != "", ImpactedRace != "Unknown")

nyc_complaniant_distribution <- nyc_complaniant %>%
  group_by(OfficerID, ComplaintID, OfficerRace, ImpactedRace) %>%
  summarize() %>%
  ungroup

nyc_complaniant_distribution <-  nyc_complaniant_distribution %>%
  group_by(OfficerRace, ImpactedRace) %>%
  summarize(num_of_complaints = n())


nyc_complaniant_distribution %>%
  ggplot() +
  geom_bar(aes(x = ImpactedRace, y = num_of_complaints), stat = "identity", fill = "brown") +
  facet_wrap(~ OfficerRace, scales = "free") +
  labs(title = "Complaints by Officer Race and Complaniant Race") +
  xlab("Complaniant Race") +
  ylab("Number of Complaints") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())

```

#### Chicago

```{r Chicago-Distribution-of-Complainants-by-Officer-Race, fig.align = 'center'}
chicago_complaniant <- read.csv("data/chicago-complaints-complainants.csv") %>%
  mutate(cr_id = as.character(cr_id)) %>%
  select(cr_id, race)

chicago_complaniant <- rename(chicago_complaniant, impacted_race = race)

chicago_complaints_and_officer <- rename(chicago_complaints_and_officer, officer_race = race)

chicago_complaniant <- chicago_complaints_and_officer %>%
  left_join(chicago_complaniant, by="cr_id") %>%
  filter(impacted_race != "", impacted_race != "Unknown")

chicago_complaniant_distribution <- chicago_complaniant %>%
  group_by(UID, cr_id, officer_race, impacted_race) %>%
  summarize() %>%
  ungroup

chicago_complaniant_distribution <-  chicago_complaniant_distribution %>%
  group_by(officer_race, impacted_race) %>%
  summarize(num_of_complaints = n())

chicago_complaniant_distribution$impacted_race[chicago_complaniant_distribution$impacted_race == "ASIAN/PACIFIC ISLANDER"] <- "ASIAN"
chicago_complaniant_distribution$officer_race[chicago_complaniant_distribution$officer_race == "ASIAN/PACIFIC ISLANDER"] <- "ASIAN"


chicago_complaniant_distribution$impacted_race[chicago_complaniant_distribution$impacted_race == "NATIVE AMERICAN/ALASKAN NATIVE"] <- "NATIVE"
chicago_complaniant_distribution$officer_race[chicago_complaniant_distribution$officer_race == "NATIVE AMERICAN/ALASKAN NATIVE"] <- "NATIVE"


chicago_complaniant_distribution %>%
  ggplot() +
  geom_bar(aes(x = impacted_race, y = num_of_complaints), stat = "identity", fill = "brown") +
  facet_wrap(~ officer_race, scales = "free") +
  labs(title = "Complaints by Officer Race and Complaniant Race") +
  xlab("Complaniant Race") +
  ylab("Number of Complaints") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  theme(panel.background = element_rect(fill = "white", colour = "white",
                                size = 2, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "gray"),
          panel.grid.major.x = element_blank())

```


